Supplemental Information A: Full Annotated Analysis and Reproducible Code

### 2022-06-28 Supplemental-Material-A-Annotated-Analysis.Rmd
MI & 3-way ANOVA daily weekly added

```{r MiMain3mpca}
library(Amelia)

varNamInMiMain <- varNames %>%
  filter(
    aux == 1,
    studyS1 == "S1",
    studyS2 == "S2",
    studyS3 == "S3"
  ) %>%
  select(varNam) %>%
  pull

@@

# MiMain <-
#   amelia(
#     dtMiMain,
#     m = 1, # 20
#     ts = "TIDnum",
#     cs = "ID",
#     idvars = c("TID", "date", "week", "study"), #"PID",
#     #lags = names(dtMiMain)[!names(dtMiMain) %in% idVars],
#     #lags = names(dtMiMain)[names(dtMiMain) %in%varNamInMiMainRed],
#     #noms = names(dtMiMain)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMain))],
#     parallel = "multicore",
#     ncpus = parallel::detectCores(),
#     p2s = 1
#   )
# save(MiMain, file = "data/imputed/MainImputed.RData")
load("data/imputed/MainImputed.RData")
```

@@

dtPcaS123Imp <- MiMain$imputations[[1]] %>% 
  select(ID, date, week, study, any_of(varNamPcaS123)) %>%
  arrange(ID, TIDnum) %>%
  filter(TIDnum <= min(max(dtS1Red$TIDnum), max(dtS2Red$TIDnum), max(dtS3Red$TIDnum)))

@@

@@

```{r threeWayAnova, message=FALSE, warning=FALSE}
library(ThreeWay)
# threeWayAno <- threewayanova(
#   Y = dtMainWide,
#   n = nrow(dtMainWide), # Nr. ppt.s (A)
#   m = length(unique(dtPcaS123$TIDnum)), # Nr. time points (B)
#   #p = sum(!varNamPcaS123 %in% idVars) # Nr. variables (C)
#   p = length(names(dtPcaS123)[!names(dtPcaS123) %in% idVars]) # Nr. variables (C)
# )

capture.output(
  threeWayAnoImp <- threewayanova(
    Y = dtMainImpWide,
    n = nrow(dtMainImpWide), # Nr. ppt.s (A)
    m = length(unique(dtPcaS123Imp$TIDnum)), # Nr. time points (B)
    p = length(names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars]) # Nr. variables (C)
  ), 
file='NUL')

options(knitr.kable.NA = '')
data.frame(SS = sapply(threeWayAnoImp, c)) %>%
  rownames_to_column(., var = "EffectNam") %>%
  mutate(RSqrd = SS / sum(SS) * 100) %>%
  add_row(
    EffectNam = "Total", 
    SS = sum(.$SS),
    RSqrd = NA
  ) %>%
  mutate(
    Effect = c(
      "Person",
      "Time",
      "Variable",
      "Person*Time",
      "Person*Variable",
      "Time*Variable",
      "Person*Time*Variable+error",
      "Total"
    )
  ) %>%
  select(Effect, everything(), -EffectNam) %>%
  kbl(.,
      escape = FALSE,
      booktabs = T,
      align = c("l", "r", "c"),
      digits=2,
      col.names = c(
        "Effect",
        "Sum of Squares",
        "$R^2$"
      ),
      caption = "Three-way ANOVA Main Analysis Single Imputation") %>%
  kable_classic(
    full_width = F,
    lightable_options = "hover",
    html_font = "Cambria"
  )

We find that the $Person*Time*Variable+error$ sums of squares are indeed the largest and speak towards the utility of a Three-Mode PCA. Interestingly, time itself has a relatively small percentage of explained variance and only becomes important in compination with person- or person- and variable-specific differences.

```{r panel-setup, include = FALSE}
xaringanExtra::use_panelset()
xaringanExtra::style_panelset_tabs(font_family = "inherit")
```

>>>>

```{r MiMain}
library(Amelia)
varNamInMiMain <- varNames %>%
  filter(
    aux == 1,
    studyS1 == "S1",
    studyS2 == "S2",
    studyS3 == "S3"
  ) %>%
  select(varNam) %>%
  pull

@@

dtMiMainDaily <- rbind(
  dtS1Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S1") %>% 
    mutate(across(!TID & !study, as.numeric)),
  dtS2Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S2"),
  dtS3Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S3")
) %>%
  group_by(study, PID) %>%
  mutate(ID = cur_group_id()) %>%
  ungroup %>%
  mutate(
    date = as.Date(gsub(" .*", "", TID)),
    week = strftime(date, format = "%Y-W%V"),
    roomate.pre = as.numeric(roomate.pre),
    Reason.pre = as.numeric(Reason.pre),
    association.pre = as.numeric(association.pre)
  ) %>%
  group_by(ID, date, study) %>%
  summarise_if(is.numeric, mean, na.rm = TRUE) %>%
  ungroup %>%
  group_by(study) %>%
  mutate(TIDnum = as.numeric(factor(date))) %>%
  ungroup %>%
  select(ID,
         study,
         date,
         everything(),
         -PID,
         -starts_with("ResponseId")) %>%
  as.data.frame %>%
  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
  complete(., ID, TIDnum) %>% # add na if id and TID combination is missing
  group_by(ID) %>%
  mutate(across(ends_with(c(".pre", ".post")),  ~ replace_na(.x, mean(.x, na.rm = TRUE)))) %>%
  mutate(study = replace_na(study, calc_mode(study))) %>%
  ungroup %>%
  group_by(study, TIDnum) %>%
  mutate(across(c("date"),  ~ replace_na(.x, calc_mode(.x)))) %>%
  ungroup %>% 
  mutate_if(is.numeric, ~ ifelse(is.nan(.), NA, .)) %>%
  as.data.frame

dtMiMainWeeekly <- rbind(
  dtS1Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S1") %>% 
    mutate(across(!TID & !study, as.numeric)),
  dtS2Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S2"),
  dtS3Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S3")
) %>%
  group_by(study, PID) %>%
  mutate(ID = cur_group_id()) %>%
  ungroup %>%
  mutate(
    date = as.Date(gsub(" .*", "", TID)),
    week = strftime(date, format = "%Y-W%V"),
    roomate.pre = as.numeric(roomate.pre),
    Reason.pre = as.numeric(Reason.pre),
    association.pre = as.numeric(association.pre)
  ) %>%
  group_by(ID, week, study) %>%
  summarise_if(is.numeric, mean, na.rm = TRUE) %>%
  ungroup %>%
  group_by(study) %>%
  mutate(TIDnum = as.numeric(factor(week))) %>%
  ungroup %>%
  select(ID,
         study,
         week,
         everything(),
         -PID,
         -starts_with("ResponseId")) %>%
  as.data.frame %>%
  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
  complete(., ID, TIDnum) %>% # add na if id and TID combination is missing
  group_by(ID) %>%
  mutate(across(ends_with(c(".pre", ".post")),  ~ replace_na(.x, mean(.x, na.rm = TRUE)))) %>%
  mutate(study = replace_na(study, calc_mode(study))) %>%
  ungroup %>%
  group_by(study, TIDnum) %>%
  mutate(across(c("week"),  ~ replace_na(.x, calc_mode(.x)))) %>%
  ungroup %>% 
  mutate_if(is.numeric, ~ ifelse(is.nan(.), NA, .)) %>%
  as.data.frame

# MiMain <-
#   amelia(
#     dtMiMain,
#     m = 1, # 20
#     ts = "TIDnum",
#     cs = "ID",
#     idvars = c("TID", "date", "week", "study"), #"PID",
#     #lags = names(dtMiMain)[!names(dtMiMain) %in% idVars],
#     #lags = names(dtMiMain)[names(dtMiMain) %in%varNamInMiMainRed],
#     #noms = names(dtMiMain)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMain))],
#     parallel = "multicore",
#     ncpus = parallel::detectCores(),
#     p2s = 1
#   )
# save(MiMain, file = "data/imputed/MainImputed.RData")
load("data/imputed/MainImputed.RData")

# MiMainDaily <-
#   amelia(
#     dtMiMainDaily,
#     m = 1, # 20
#     ts = "TIDnum",
#     cs = "ID",
#     idvars = c("date", "study"), #"PID",
#     #lags = names(dtMiMainDaily)[!names(dtMiMainDaily) %in% idVars],
#     #lags = names(dtMiMainDaily)[names(dtMiMainDaily) %in%varNamInMiMainRed],
#     #noms = names(dtMiMainDaily)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMainDaily))],
#     parallel = "multicore",
#     ncpus = parallel::detectCores(),
#     p2s = 1
#   )
# save(MiMainDaily, file = "data/imputed/MainDailyImputed.RData")
load("data/imputed/MainDailyImputed.RData")

# MiMainWeekly <-
#   amelia(
#     dtMiMainWeeekly,
#     m = 1, # 20
#     ts = "TIDnum",
#     cs = "ID",
#     idvars = c("week", "study"), #"PID",
#     #lags = names(dtMiMainWeeekly)[!names(dtMiMainWeeekly) %in% idVars],
#     #lags = names(dtMiMainWeeekly)[names(dtMiMainWeeekly) %in%varNamInMiMainRed],
#     #noms = names(dtMiMainWeeekly)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMainWeeekly))],
#     parallel = "multicore",
#     ncpus = parallel::detectCores(),
#     p2s = 1
#   )
# save(MiMainWeekly, file = "data/imputed/MainWeeklyImputed.RData")
load("data/imputed/MainWeeklyImputed.RData")
```

@@

dtPcaS123Imp <- MiMain$imputations[[1]] %>% 
  select(ID, date, week, study, any_of(varNamPcaS123)) %>%
  arrange(ID, TIDnum) %>%
  filter(TIDnum <= min(max(dtS1Red$TIDnum), max(dtS2Red$TIDnum), max(dtS3Red$TIDnum)))
dtPcaS123DailyImp <- MiMainDaily$imputations[[1]] %>%
  select(ID, date, study, any_of(varNamPcaS123)) %>%
  arrange(ID, TIDnum) %>%
  filter(
    TIDnum <= dtMiMainDaily %>% group_by(study) %>% summarise(TIDmax = max(TIDnum)) %>% select(TIDmax) %>% min
  )
dtPcaS123WeeklyImp <- MiMainWeekly$imputations[[1]] %>%
  select(ID, week, study, any_of(varNamPcaS123)) %>%
  arrange(ID, TIDnum) %>%
  filter(
    TIDnum <= dtMiMainWeeekly %>% group_by(study) %>% summarise(TIDmax = max(TIDnum)) %>% select(TIDmax) %>% min
  )

@@

dtMainDailyImpWide <- dtPcaS123DailyImp %>%
  select(
    -date
  ) %>% 
  melt(
    .,
    id=c("ID", "TIDnum", "study")
  ) %>%
  mutate(
    variable = paste(variable, stringr::str_pad(TIDnum, width = 2, pad = 0), sep = "_")
  ) %>%
  select(-TIDnum, -study) %>%
  arrange(variable, ID) %>%
  pivot_wider(names_from = variable, values_from = value) %>%
  #replace(is.na(.), 0) %>%
  select(-ID)

dtMainWeeklyImpWide <- dtPcaS123WeeklyImp %>%
  select(
    -week
  ) %>% 
  melt(
    .,
    id=c("ID", "TIDnum", "study")
  ) %>%
  mutate(
    variable = paste(variable, stringr::str_pad(TIDnum, width = 2, pad = 0), sep = "_")
  ) %>%
  select(-TIDnum, -study) %>%
  arrange(variable, ID) %>%
  pivot_wider(names_from = variable, values_from = value) %>%
  #replace(is.na(.), 0) %>%
  select(-ID)
```

@@

```{r threeWayAnovaMainBiDaily, message=FALSE, warning=FALSE}
library(ThreeWay)
capture.output(
  threeWayAnoBiDailyImp <- threewayanova(
    Y = dtMainImpWide,
    n = nrow(dtMainImpWide), # Nr. ppt.s (A)
    m = length(unique(dtPcaS123Imp$TIDnum)), # Nr. time points (B)
    p = length(names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars]) # Nr. variables (C)
  ), 
file='NUL')

options(knitr.kable.NA = '')
data.frame(SS = sapply(threeWayAnoBiDailyImp, c)) %>%
  rownames_to_column(., var = "EffectNam") %>%
  mutate(RSqrd = SS / sum(SS) * 100) %>%
  add_row(
    EffectNam = "Total", 
    SS = sum(.$SS),
    RSqrd = NA
  ) %>%
  mutate(
    Effect = c(
      "Person",
      "Time",
      "Variable",
      "Person*Time",
      "Person*Variable",
      "Time*Variable",
      "Person*Time*Variable+error",
      "Total"
    )
  ) %>%
  select(Effect, everything(), -EffectNam) %>%
  kbl(.,
      escape = FALSE,
      booktabs = T,
      align = c("l", "r", "c"),
      digits=2,
      col.names = c(
        "Effect",
        "Sum of Squares",
        "$R^2$"
      ),
      caption = "Three-way ANOVA Main Analysis Bi-daily (Single Imputation)") %>%
  kable_classic(
    full_width = F,
    lightable_options = "hover",
    html_font = "Cambria"
  )
```

<br>

We find that for the bi-daily time scale the $Person*Time*Variable+error$ sums of squares are indeed the largest and speak towards the utility of a Three-Mode PCA. Interestingly, time itself has a relatively small percentage of explained variance and only becomes important in compination with person- or person- and variable-specific differences.

```{r threeWayAnovaMainDaily, message=FALSE, warning=FALSE}
capture.output(
  threeWayAnoDailyImp <- threewayanova(
    Y = dtMainDailyImpWide,
    n = nrow(dtMainDailyImpWide), # Nr. ppt.s (A)
    m = length(unique(dtPcaS123DailyImp$TIDnum)), # Nr. time points (B)
    p = length(names(dtPcaS123DailyImp)[!names(dtPcaS123DailyImp) %in% idVars]) # Nr. variables (C)
  ), 
file='NUL')

options(knitr.kable.NA = '')
data.frame(SS = sapply(threeWayAnoDailyImp, c)) %>%
  rownames_to_column(., var = "EffectNam") %>%
  mutate(RSqrd = SS / sum(SS) * 100) %>%
  add_row(
    EffectNam = "Total", 
    SS = sum(.$SS),
    RSqrd = NA
  ) %>%
  mutate(
    Effect = c(
      "Person",
      "Time",
      "Variable",
      "Person*Time",
      "Person*Variable",
      "Time*Variable",
      "Person*Time*Variable+error",
      "Total"
    )
  ) %>%
  select(Effect, everything(), -EffectNam) %>%
  kbl(.,
      escape = FALSE,
      booktabs = T,
      align = c("l", "r", "c"),
      digits=2,
      col.names = c(
        "Effect",
        "Sum of Squares",
        "$R^2$"
      ),
      caption = "Three-way ANOVA Main Analysis Daily (Single Imputation)") %>%
  kable_classic(
    full_width = F,
    lightable_options = "hover",
    html_font = "Cambria"
  )
```

<br>

We additionally see that for the daily time scale aggregation, the $Person*Time*Variable+error$ sums of squares are reduced but remain the largest sums of squares and are still substantial.

```{r threeWayAnovaMainWweekly, message=FALSE, warning=FALSE}
capture.output(
  threeWayAnoWeeklyImp <- threewayanova(
    Y = dtMainWeeklyImpWide,
    n = nrow(dtMainWeeklyImpWide), # Nr. ppt.s (A)
    m = length(unique(dtPcaS123WeeklyImp$TIDnum)), # Nr. time points (B)
    p = length(names(dtPcaS123WeeklyImp)[!names(dtPcaS123WeeklyImp) %in% idVars]) # Nr. variables (C)
  ), 
file='NUL')

options(knitr.kable.NA = '')
data.frame(SS = sapply(threeWayAnoWeeklyImp, c)) %>%
  rownames_to_column(., var = "EffectNam") %>%
  mutate(RSqrd = SS / sum(SS) * 100) %>%
  add_row(
    EffectNam = "Total", 
    SS = sum(.$SS),
    RSqrd = NA
  ) %>%
  mutate(
    Effect = c(
      "Person",
      "Time",
      "Variable",
      "Person*Time",
      "Person*Variable",
      "Time*Variable",
      "Person*Time*Variable+error",
      "Total"
    )
  ) %>%
  select(Effect, everything(), -EffectNam) %>%
  kbl(.,
      escape = FALSE,
      booktabs = T,
      align = c("l", "r", "c"),
      digits=2,
      col.names = c(
        "Effect",
        "Sum of Squares",
        "$R^2$"
      ),
      caption = "Three-way ANOVA Main Analysis Weekly (Single Imputation)") %>%
  kable_classic(
    full_width = F,
    lightable_options = "hover",
    html_font = "Cambria"
  )
```

<br>

However, with the aggregation to `r length(unique(dtPcaS123WeeklyImp$TIDnum))` weekly measurements, the $Person*Time*Variable+error$ sums of squares become less prominent and the variable differences explain substantially more variance.
